---
title: "PSYC 245 Exam Formulas and Tables"
subtitle: "(scroll down for the unit you need)"
author: "Patrick Bolger, PhD"
date: "`r format(Sys.time(), '%d %B %Y')`"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  rmdformats::downcute:
    use_bookdown: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, comment = NA )
```

```{r pload, echo = F}

pacman::p_load(readr, DT, dplyr)

```

# Exam 2

## The Interquartile Range (*IQR*)

$$Q_3-Q_1$$
<br/>

## The mean

$$\overline{x}=\frac{\sum{x_i}}{n}$$
<br/>

## The variance


$$s^2=\frac{\sum(x_i-\overline{x})^2}{n-1}$$

<br/>

## the standard deviation

$$s=\sqrt{\frac{\sum(x_i-\overline{x})^2}{n-1}}$$
<br/>

## the *z*-score


$$z_i=\frac{x_i-\overline{x}}{s}$$

<br/>

## a raw score, given a z-score, an SD, and a mean

$$x_i=z_i(s)+\overline{x}$$

<br/>

## Pearson's *r* correlation coefficient

$$r_{x,y}=\frac{\sum z_{(x_i)}z_{(y_i)}}{n-1}$$

<br/>

## Simple Regression equation

$$y=a+bX$$

## Normal (*z*-) tables

```{r importZ, include=F}

# raw file shows area to the left of specific z-scores
ztable_leftward <- round(read_csv("tables/ztable.csv", col_types = cols()),4)
# remove lines that are all zeros
ztable_leftward_filtered <- ztable_leftward %>% 
  filter(if_all(starts_with("0"), ~ . > 0)) %>%
  filter(if_all(starts_with("0"), ~ . < 1)) 
ztable_rightward_filtered <- ztable_leftward_filtered %>% mutate(across(`0`:`0.09`, ~ 1-.x))

ztable_leftward_filtered <- ztable_leftward_filtered %>% mutate(across(`0`:`0.09`, ~ round(.x,4)))
ztable_rightward_filtered <- ztable_rightward_filtered %>% mutate(across(`0`:`0.09`, ~ round(.x,4)))

```

```{r zContainer}

zsketch <- htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(rowspan = 2, 'First 2 digits of z-score'),
      th(colspan = 10, '3rd digit of z-score'),
    ),
    tr(lapply(
      c('0.00', '0.01', '0.02', '0.03', '0.04', 
        '0.05', '0.06', '0.07', '0.08', '0.09'), th)
      )
    )
  )
)

```

### *z*-table (rightward)

**Directions**: To get any particular *z*-score, combine the value in the row in the left-most column (`First 2 digits of z-score`) with the appropriate value across the top of the rightward columns (`3rd digit of z-score`). For example, the area `to the right` of a *z*-score of `-1.25` would be found by finding the intersection of `-1.2` in the first column with the column labeled `0.05` across the top (This value is 0.8944).

<br/>

```{r zDisplayRightward}

DT::datatable(
  ztable_rightward_filtered, 
  container = zsketch,
  rownames = FALSE,
  extensions = 'KeyTable',
  options = list(keys = TRUE)
  )

```
<br/>

### *z*-table (leftward)

**Directions**: To get any particular *z*-score, combine the value in the row in the left-most column (`First 2 digits of z-score`) with the appropriate value across the top of the rightward columns (`3rd digit of z-score`). For example, the area `to the left` of a *z*-score of `-1.25` would be found by finding the intersection of `-1.2` in the first column with the column labeled `0.05` across the top (this value is 0.1056).

<br/>

```{r zDisplayLeftward}

DT::datatable(
  ztable_leftward_filtered, 
  container = zsketch,
  rownames = FALSE,
  extensions = 'KeyTable',
  options = list(keys = TRUE),
  caption = 
  )

```

---


# Exam 3

## Odds

$$Odds = A\ to\ B$$

## probability from odds

$$\frac{B}{A+B}$$

## expected value

$$Expected\ value=a_1p_1+a_2p_2+\ ...a_kp_k$$

, where

*a*: possible outcomes  
*p*: associated probabilities  
*k*: number of possible outcomes  

---

<br/>

# Exam 4

## **Rough** 95% Confidence Interval for a proportion

$$95\%\ CI_p= \hat{p}\pm\Bigg(2\times\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}\Bigg)$$
<br/>

## **More exact** Confidence Intervals for a proportion

$$CI_p= \hat{p}\pm\Bigg(z^*\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}\Bigg)$$
<br/>

## Important $z^*$ values

- 95th percentile  
  - 1.64  
<br/>
- Middle 95% of normal curve  
  - [-1.96, +1.96]
- 97.5th percentile  
  - 1.96  
- 2.5th percentile  
  - -1.96  
<br/>
- Middle 99% of normal curve  
  - [-2.58, +2.58]  
- 99.5th percentile  
  - 2.58  
- 0.5th percentile  
  - -2.58  
<br/>
(for more, see [http://www.normaltable.com/](http://www.normaltable.com/))
  
<br/>

## Confidence Intervals for a mean

$$95\%\ CI_{\mu}= \bar{x}\pm\Bigg(z^*\times\frac{s}{\sqrt{n}}\Bigg)$$
<br/>

## *p*-value

$$p=P\big(D|H_0\big)$$
<br/>

## Test statistics

<br/>

---

### *z*-test for a proportion

$$z=\frac{\hat{p}-p_{null}}{SE_{\hat{p}}}$$
<br/>

[the sample proportion ($\hat{p}$) minus the proportion assumed under the null hypothesis ($p_{null}$), divided by the margin of error (*MoE*) of the proportion ($SE_{\hat{p}}$)]

<br/>

#### standard error for a sample proportion

$$SE_{\hat{p}}=\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}$$
<br/>

#### Combined formula for a *z*-test for a proportion

$$z=\frac{\hat{p}-p_{null}}{\sqrt{\frac{\hat{p}(1-\hat{p})}{n}}}$$

---

### *z*-test for a mean

$$z=\frac{\bar{x}-\mu_{null}}{SE_{mean}}$$
<br/>

[the sample mean ($\bar{x}$) minus the mean assumed under the null hypothesis ($\mu_{null}$), divided by the standard error of the mean ($SE_{mean}$)]

<br/>

#### The standard error of the mean

$$SE_{mean}=\frac{s}{\sqrt{n}}$$
<br/>

#### Combined formula for a *z*-test of a mean

$$z=\frac{\bar{x}-\mu_{null}}{s \big/\sqrt{n}}$$